The Weakly Nonlinear Magnetorotational Instability in a Taylor Couette Flow

### 2016-07-13 paper/widegap_mri_paper.tex
all helical terms added
added fiducial parameters table to widegap paper. unfinished. do we like this?

%The MRI is fundamental to our understanding of star formation, but its

@@

Thus through judicious choice of cylinder rotation rates, we can set $q(r_0) = 3/2$, for quasi-Keplerian flow. Note that in the narrow gap approximation we have the freedom to choose any rotation profile, as the narrow gap width imposes a linear shear (constant $q$). Our base velocity is $\mathbf{u_0} = r\Omega(r) \phihat$. We initialize an axial magnetic field $\mathbf{B_0} = B_0 \zhat$.

@@

\beq
\begin{split}
\label{eq:Psi_perturbed}
\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] \\
= - J(\Psi, \frac{1}{r^2} ( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi) ) + \frac{2}{\beta} J(A, \frac{1}{r^2} ( \nabla^2 A - \frac{2}{r} \partial_rA) ) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi \\
\end{split}
\eeq

\begin{align}
& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = \frac{2}{\beta} \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \frac{2}{\beta}\frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi \label{eq:uphi_perturbed} \\
& \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} ( \nabla^2 A - \frac{2}{r} \partial_r A )= \frac{1}{r} J(A, \psi)  \label{eq:A_perturbed}\\
  \label{eq:Bphi_perturbed}
& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)
+ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A 
\end{align}

@@

\beq\label{eq:unperturbed_matrix_eqns}
\mathcal{D}\partial_t\mathbf{V} + \mathcal{L}\mathbf{V} + \epsilon^2 \widetilde{\mathcal{G}} \mathbf{V} + \mathbf{N} = 0,
\eeq

@@

\begin{align}
\mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3. \label{eq:ordere3}
\end{align}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
\caption{Eigenfunctions}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
\caption{Nonlinear terms}
\end{figure}

>>>>

%The MRI is fundamental to our understanding of star formation, but its 
- people make a number of simplifications, incl thin gap, shearing box, etc

@@

Thus through judicious choice of cylinder rotation rates, we can set $q(r_0) = 3/2$, for quasi-Keplerian flow. Note that in the narrow gap approximation we have the freedom to choose any rotation profile, as the narrow gap width imposes a linear shear (constant $q$). Our base velocity is $\mathbf{u_0} = r\Omega(r) \phihat$. We initialize a magnetic field $\mathbf{B_0} = B_0 \zhat + B_0 \xi \frac{r_0}{r} \phihat$, so that the base magnetic field is axial when $\xi = 0$ and otherwise helical.

In this work we will focus our findings on two fiducial parameter sets, one for the standard MRI (hereafter SMRI) where $\xi = 0$ and one for the helical MRI (hereafter HMRI). Our fiducial parameters are described in Table \ref{table:parameters}.

\begin{table}[ht]\label{table:parameters}
\normalsize
\caption{Fiducial parameters for MRI runs} 
\centering
%\resizebox{\textwidth}{!}
{\begin{tabular}{rlrrrrrrr}
  \hline
 & $\xi$ & $Pm$ & $\Omega_2/\Omega_1$ & $\beta$ & X \\ 
  \hline\hline
Standard MRI & 0 & 25 & 0 & 0 & 0 \\ 
Helical MRI & 4 & 0 & 0\\ 
   \hline
\end{tabular}}
\end{table}

@@

\beq
\begin{split}
\label{eq:Psi_perturbed}
\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] \\
= - J(\Psi, \frac{1}{r^2} ( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi) ) + \frac{2}{\beta} J(A, \frac{1}{r^2} ( \nabla^2 A - \frac{2}{r} \partial_rA) ) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi \\
\end{split}
\eeq

\begin{align}
& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = \frac{2}{\beta} \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \frac{2}{\beta}\frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi \label{eq:uphi_perturbed} \\
& \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} ( \nabla^2 A - \frac{2}{r} \partial_r A )= \frac{1}{r} J(A, \psi)  \label{eq:A_perturbed}\\
  \label{eq:Bphi_perturbed}
& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{2}{r^3} B_0 \xi \partial_z \Psi - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)
+ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A 
\end{align}

@@

\beq\label{eq:unperturbed_matrix_eqns}
\mathcal{D}\partial_t\mathbf{V} + \mathcal{L}\mathbf{V} + \epsilon^2 \widetilde{\mathcal{G}} \mathbf{V} + \xi \widetilde{\mathcal{H}} \mathbf{V} +  \mathbf{N} = 0,
\eeq

@@

\begin{align}
\mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 + \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0. \label{eq:ordere3}
\end{align}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
\caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations. We use our fiducial parameters for the standard MRI ($\xi = 0$). Eigenfunctions are normalized such that either the real or imaginary part is zero.}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}
\end{figure}
